Kalman Filters: Derivation

### 2015-07-26 ms-thesis/ver2/kalman/ms-thesis-kalman.tex
-

Kalman Filters, or in general, SSM methods try to model dynamic data using
two-level graphic model. In the first level, there are hidden states, one per
time step, which are random variables distributed as Gaussians. An outside
observer (modeler) does not see these states directly, instead sees them through
an ``observed'' variable whose outputs are assumed to be distorted by another
Gaussian distribution, refered to as white noise. In each time step, it is
assumed the hidden variables evolve linearly, in linear algebra this means the
hidden state is multiplied by a constant matrix $A$. Mathematically,
\begin{eqnarray*}
x_{t+1} &=& Ax_t + v_t 
\end{eqnarray*}
where
\begin{itemize}
   \item $v_0,v_1,...$ is white Gaussian noise with $\mathbf{E}v_t = 0$ and $cov
   \: v_t = Q$
   \item $x_t \sim N(\mu_t, \Sigma_t)$.
\end{itemize}
In order to model the output variables, we multiply hidden states with another
constant matrix and add another Gaussian white noise. All together our SSM looks
as follows:
\begin{eqnarray}
x_{t+1} = Ax_t + v_t \label{m1}\\
y_t = Cx_t + w_t \label{m2}
\end{eqnarray}
where $w_0,w_1,...$ are white Gaussian noise with $\mathbf{E}w_t = 0$ and $cov
\:\: w_t = R$. Random variables $w_t$ and $v_t$ are independent of $x_t$
and $y_t$, and from eachother, this assumption is put forth for simplicity. The
dimensions of each variable are:

@@

\begin{figure}[!hbp]
\center{
  \scalebox{0.6}{
  \includegraphics{./kalman/kf.JPG}
  }
}
\caption{Hidden and observed states of a KF}
\end{figure}

@@

\begin{itemize}
   \item $\hat{x}_{t}^t \triangleq \mathbf{E}[x_t|y_0,...,y_t]$
   \item $P_{t}^t \triangleq \mathbf{E}[(x_t - \hat{x}_{t|t})(x_t -
   \hat{x}_{t|t})'|y_0,...,y_t]$ 
\end{itemize}
Our first goal is going to be trying to calculate $\hat{x}_{t+1}^{t+1}$ and
$P_{t+1}^{t+1}$ using a new measurement $y_{t+1}$. For this, we need to
``reverse the arrow'' so to speak, in other words formulize the distribution
where $x_t$ has conditional dependence on $y_t$ instead of other way around,
which is what the base equations state right now.

If we take expectation of both sides of \eqref{m1}
\begin{eqnarray*}
\mathbf{E}x_{t+1} = \hat{x}_{t+1} = A\mu_t = A\hat{x}_t
\end{eqnarray*}
Taking covariance of both sides of \eqref{m1} and denoting $P_t$ as $cov \: x(t)$
\begin{eqnarray*}
P_{t+1} = AP_{t}A' + Q
\end{eqnarray*}
This transition is called a ``time update''. It will allow us to propagate the
normal distribution at time $t$ to time $t_1$. The formulas that contain given
$y$'s use the same technique;
\begin{eqnarray*}
\hat{x}_{t+1}^t = Ax_{t}^t\\
P_{t+1}^t = AP_{t}^tA' + Q
\end{eqnarray*}
\begin{eqnarray*}
y_{t+1} &=& Cx_{t+1} + w_t\\
\mathbf{E}[y_{t+1}|y_0,...,y_t] &=& \mathbf{E}[Cx_{t+1}+w_t|y_0,...,y_t]\\
\hat{y}_{t+1}^t &=& C\hat{x}_{t+1}
\end{eqnarray*}
Similarly for covariance
\begin{eqnarray*}
E[(y_{t+1}-\hat{y}_{t+1}^t)(y_{t+1}-\hat{y}_{t+1}^t)'|y_0,...,y_t] =
CP_{t+1}^tC' + R
\end{eqnarray*}

@@

In order to define the joint normal distribution of $x_{t}$ and $y_{t}$, we need
to write the joint mean and covariance of this big Gaussian. Writing down the
mean will be easy, covariance will be little harder to derive. We can use a
trick \cite{lall} where we write $y_{t} = Cx_{t} + w_t$ as $z = Hu$
where
\begin{eqnarray*}
z =   \left[ \begin{array}{c}
      x_{t} \\
      y_{t}
      \end{array} \right],
H =   \left[ \begin{array}{cc}
      I & 0 \\
      C & I
      \end{array} \right]
u =   \left[ \begin{array}{c}
      x_{t} \\
      w_t
      \end{array} \right],      
\end{eqnarray*}
Now we need to take the covariance of the simpler equation:
\begin{eqnarray*}
cov(z) &=& H cov(u) H'\\
cov(u) &=& \left[ \begin{array}{cc}
      P_{t} & 0 \\
      0 & R
      \end{array} \right]     
\end{eqnarray*}
The full product is:
\begin{eqnarray*}
\left[ \begin{array}{cc}
      I & 0 \\
      C & I
      \end{array} \right]
\left[ \begin{array}{cc}
      P_{t} & 0 \\
      0 & R
      \end{array} \right]
\left[ \begin{array}{cc}
      I & C' \\
      0 & I
      \end{array} \right]           
\end{eqnarray*}
which results in
\begin{eqnarray*}
\left[ \begin{array}{cc}
      P_{t} & P_{t}C' \\
      CP_{t} & CP_{t}C'+R
      \end{array} \right]           
\end{eqnarray*}
We can also re-write this for conditional and with the mean;
\begin{eqnarray}
\left[ \begin{array}{c}
      \hat{x}_{t}^t \\
      C\hat{x}_{t}^t
      \end{array} \right]
\:and\:      
\left[ \begin{array}{cc}
      P_{t}^t & P_{t}^tC' \\
      CP_{t}^t & CP_{t}^tC'+R
      \end{array} \right] \label{jointxy}
\end{eqnarray}
Same for the joint distribution for $x_{t+1}, y_{t+1}$.
\begin{eqnarray}
\left[ \begin{array}{c}
      \hat{x}_{t+1}^t \\
      C\hat{x}_{t+1}^t
      \end{array} \right]
\:and\:      
\left[ \begin{array}{cc}
      P_{t+1}^t & P_{t+1}^tC' \\
      CP_{t+1}^t & CP_{t+1}^tC'+R
      \end{array} \right] \label{jointxy2}
\end{eqnarray}

Now in order to get statements for mean and variance for $x_{t+1}^{t+1}$, we
need to understand partitioned Gaussians. This will guide us when we use
portions of \eqref{jointxy2} and deriving the final equation \cite{jordan}.

@@

\begin{eqnarray}
\mu = 
\left[ \begin{array}{c}
      \mu_1 \\
      \mu_2
      \end{array} \right]           
\Sigma = 
\left[ \begin{array}{cc}
      \Sigma_{11} & \Sigma_{12}\\
      \Sigma_{21} & \Sigma_{22}
      \end{array} \right] \label{condgeneric}
\end{eqnarray}
\begin{eqnarray*}
p(x|\mu,\Sigma) &=& \frac{1}{(2\pi)^{(p+q)/2}|\Sigma|^{1/2}} \\
&& exp\bigg\{
-\frac{1}{2}
\left( \begin{array}{c}
      x_1 - \mu_1 \\
      x_2 - \mu_2
      \end{array} \right)'
\bigg\}
\left[ \begin{array}{cc}
      \Sigma_{11} & \Sigma_{12}\\
      \Sigma_{21} & \Sigma_{22}
      \end{array} \right]^{-1}
\left( \begin{array}{c}
      x_1 - \mu_1 \\
      x_2 - \mu_2
      \end{array} \right)
\bigg\} 
\end{eqnarray*}
After much algebra, we can obtain equation for $p(x_1|x_2)$. From this, we can
get {\em conditioned} $\mu$ and $\Sigma$, as
\begin{eqnarray}
\mu_{1|2} &=& \mu_1 + \Sigma_{12}\Sigma_{22}^{-1}(x_2 - \mu_2) \label{condx} \\
\Sigma_{1|2} &=& \Sigma_{11}-\Sigma_{12}\Sigma_{22}^{-1}\Sigma_{21} \nonumber
\end{eqnarray}
Now plugging elements from  \eqref{condx} into \eqref{jointxy2} according
to placement in \eqref{condgeneric} we can formulize $\hat{x}_{t+1}^{t+1}$ and
$P_{t+1}^{t+1}$. 
\begin{eqnarray*}
\hat{x}_{t+1}^{t+1} &=& \hat{x}_{t+1}^t + P_{t+1}^tC'
                  (CP_{t+1}^tC'+ \Sigma_{w})^{-1} \\
               &&   (y_{t+1} - C\hat{x}_{t+1}^t) \\
P_{t+1}^{t+1} &=& P_{t+1}^t - P_{t+1}^tC' (CP_{t+1}^tC' + R)^{-1}CP_{t+1}^t
\end{eqnarray*}
If we declare $K_t \triangleq P_{t+1}^tC' (CP_{t+1}^tC'+ \Sigma_{w})^{-1}$ then,
\begin{eqnarray}
\hat{x}_{t+1}^{t+1} &=& \hat{x}_{t+1}^t + K_t (y_{t+1} -
C\hat{x}_{t+1}^t) \label{xtplus1} \\
P_{t+1}^{t+1} &=& P_{t+1}^t - K_t C P_{t+1}^t
\end{eqnarray}

@@

Based on \eqref{jointxy}, we can write the covariance of joint distribution
$(x_t,x_{t+1})$. In \eqref{jointxy} the multiplication was with $C$, here it's
$A$. Hence,
\begin{eqnarray*}
\left[ \begin{array}{cc}
      P_{t}^t & P_{t}^tA' \\
      AP_{t}^t & AP_{t}^tA'+R
      \end{array} \right]
\end{eqnarray*}
Combining the previous equations together, we need to reach to a statement for
$x_{t}'$. We can start with $\mathbf{E}[x_t|x_{t+1},y_0,...,y_t)]$.
\begin{eqnarray*}
\mu_{1|2} &=& \mu_1 + \Sigma_{12}\Sigma_{22}^{-1}(x_2 - \mu_2)
\end{eqnarray*}
which becomes
\begin{eqnarray*}
\mathbf{E}[x_t|x_{t+1},y_0,...,y_t)] = \hat{x}_{t}^t+P_{t}^tA'(P_{t+1}^t)^{-1}(x_{t+1}-\hat{x}_{t+1}^t)
\end{eqnarray*}
Define $L_t \triangleq P_{t}^tA'(P_{t+1}^t)^{-1}$. Then,
\begin{eqnarray*}
\mathbf{E}[x_t|x_{t+1},y_0,...,y_t)] =
\hat{x}_{t}^t+L_t(x_{t+1}-\hat{x}_{t+1}^t)
\end{eqnarray*}
How about $Var[x_t|x_{t+1},y_0,...,y_t]$? If we remember the joint covariance
for $(x_t,x_{t+1})$
\begin{eqnarray*}
 \left[ \begin{array}{cc} 
      P_{t}^t & P_{t}^tA' \\
      AP_{t}^t & AP_{t}^tA'+R
      \end{array} \right] && \\
\end{eqnarray*}
Then
\begin{eqnarray*}
\Sigma_{1|2} &=& \Sigma_{11}-\Sigma_{12}\Sigma_{22}^{-1}\Sigma_{21}
\end{eqnarray*}
becomes
\begin{eqnarray*}
Var[x_t|x_{t+1},y_0,...,y_t] &=& P_{t}^t - P_{t}^tA'(P_{t+1}^t)^{-1}AP_{t}^t
\end{eqnarray*}
Why?
\begin{eqnarray*}
\left[ \begin{array}{cc}
      P_{t}^t & P_{t}^tA' \\
      AP_{t}^t & AP_{t}^tA'+R
      \end{array} \right]
=      
\left[ \begin{array}{cc}
      P_{t}^t & P_{t}^tA' \\
      AP_{t}^t & P_{t+1}^t
      \end{array} \right]
\end{eqnarray*}
If we try to substitute $L_t \triangleq P_{t}^tA'(P_{t+1}^t)^{-1}$ in to the
equation:
\begin{eqnarray*}
Var[x_t|x_{t+1},y_0,...,y_t] &=& P_{t}^t - P_{t}^t A'(P_{t+1}^t)^{-1}AP_{t}^t \\
&=& P_{t}^t - P_{t}^tA'(P_{t+1}^t)^{-1} (P_{t+1}^t(P_{t+1}^t)^{-1}) AP_{t}^t\\
&=& P_{t}^t - (P_{t}^tA'(P_{t+1}^t)^{-1}) P_{t+1}^t ((P_{t+1}^t)^{-1} AP_{t}^t)\\
&=& P_{t}^t - L_t  P_{t+1}^t L_t'
\end{eqnarray*}
Due to conditional independence, when we condition on $x_{t+1}$, we make $x_t$
independent of future observations $y_{t+1},...,y_T$. Hence we can simply say;
\begin{eqnarray*}
E[x_t|x_{t+1},y_0,...,y_T]  &=& E[x_t|x_{t+1},y_0,...,y_t]\\
&=& \hat{x}_{t}^t +L_t(x_{t+1}-\hat{x}_{t+1}^t) \\
Var[x_t|x_{t+1},y_0,...,y_T]  &=& Var[x_t|x_{t+1},y_0,...,y_t] \\
&=& P_{t}^t - L_t  P_{t+1}^t L_t'
\end{eqnarray*}
We are almost done. If we can drop the $x_{t+1}$ from LHS of both equations, we
can have our smoothing equations. Due to the so-called Tower Property:
\begin{eqnarray*}
\mathbf{E}[X|Z] &=& \mathbf{E}[\mathbf{E}[X|Y,Z]|Z]\\
Var[X|Z] &=& Var[\mathbf{E}[X|Y,Z]|Z] + \mathbf{E}[Var[X|Y,Z]|Z]
\end{eqnarray*}
we can drop $x_{t+1}$ from LHS.
\begin{eqnarray}
\hat{x}_{t}' & \triangleq & \mathbf{E}[x_t|y_0,...,y_T] \nonumber\\
& = & \mathbf{E}[\mathbf{E}[x_t|x_{t+1},y_0,...,y_T]|y_0,...,y_T] \nonumber\\
& = & \mathbf{E}[\hat{x}_{t}^t +L_t(x_{t+1}-\hat{x}_{t+1}^t)] \nonumber\\
& = & \hat{x}_{t}^t +L_t(\hat{x}_{t+1}' -\hat{x}_{t+1}^t) \label{xsmooth}
\end{eqnarray}
Using similar algebra, we get
\begin{eqnarray*}
P_{t}'  = P_{t}^t + L_t(P_{t+1}'-P_{t+1}^t)L_t'
\end{eqnarray*}

>>>>

Kalman Filters, or in general, SSM methods try to model dynamic data using
two-level graphic model. In the first level, there are hidden states, one per
time step, which are random variables distributed as Gaussians. An outside
observer (modeler) does not see these states directly, instead sees them through
an ``observed'' variable whose outputs are assumed to be distorted by another
Gaussian distribution, refered to as white noise. In each time step, it is
assumed the hidden variables evolve linearly, in linear algebra this means the
hidden state is multiplied by a constant matrix $A$. Mathematically,
\begin{eqnarray*}
x_{t+1} &=& Ax_t + v_t 
\end{eqnarray*}
where
\begin{itemize}
   \item $v_0,v_1,...$ is white Gaussian noise with $\mathbf{E}v_t = 0$ and $cov
   \: v_t = Q$
   \item $x_t \sim N(\mu_t, \Sigma_t)$.
\end{itemize}
In order to model the output variables, we multiply hidden states with another
constant matrix and add another Gaussian white noise. All together our SSM looks
as follows:
\begin{eqnarray}
x_{t+1} = Ax_t + v_t \mlabel{1}\\
y_t = Cx_t + w_t \mlabel{2}
\end{eqnarray}
where $w_0,w_1,...$ are white Gaussian noise with $\mathbf{E}w_t = 0$ and $cov
\:\: w_t = R$. Random variables $w_t$ and $v_t$ are independent of $x_t$
and $y_t$, and from eachother, this assumption is put forth for simplicity. The
dimensions of each variable are:

@@

@@

\begin{itemize}
   \item $\hat{x}_{t}^t \equiv \mathbf{E}[x_t|y_0,...,y_t]$
   \item $P_{t}^t \equiv \mathbf{E}[(x_t - \hat{x}_{t|t})(x_t -
   \hat{x}_{t|t})'|y_0,...,y_t]$ 
\end{itemize}
Our first goal is going to be trying to calculate $\hat{x}_{t+1}^{t+1}$ and
$P_{t+1}^{t+1}$ using a new measurement $y_{t+1}$. For this, we need to
``reverse the arrow'' so to speak, in other words formulize the distribution
where $x_t$ has conditional dependence on $y_t$ instead of other way around,
which is what the base equations state right now.

If we take expectation of both sides of (1)
\begin{eqnarray*}
\mathbf{E}x_{t+1} = \hat{x}_{t+1} = A\mu_t = A\hat{x}_t
\end{eqnarray*}
Taking covariance of both sides of (1) and denoting $P_t$ as $cov \: x(t)$
\begin{eqnarray*}
P_{t+1} = AP_{t}A' + Q
\end{eqnarray*}
This transition is called a ``time update''. It will allow us to propagate the
normal distribution at time $t$ to time $t_1$. The formulas that contain given
$y$'s use the same technique;
\begin{eqnarray*}
\hat{x}_{t+1}^t = Ax_{t}^t\\
P_{t+1}^t = AP_{t}^tA' + Q
\end{eqnarray*}
\begin{eqnarray*}
y_{t+1} &=& Cx_{t+1} + w_t\\
\mathbf{E}[y_{t+1}|y_0,...,y_t] &=& \mathbf{E}[Cx_{t+1}+w_t|y_0,...,y_t]\\
\hat{y}_{t+1}^t &=& C\hat{x}_{t+1}
\end{eqnarray*}
Similarly for covariance
\begin{eqnarray*}
E[(y_{t+1}-\hat{y}_{t+1}^t)(y_{t+1}-\hat{y}_{t+1}^t)'|y_0,...,y_t] =
CP_{t+1}^tC' + R
\end{eqnarray*}

@@

In order to define the joint normal distribution of $x_{t}$ and $y_{t}$, we need
to write the joint mean and covariance of this big Gaussian. Writing down the
mean will be easy, covariance will be little harder to derive. We can use a
trick \cite{lall} where we write $y_{t} = Cx_{t} + w_t$ as $z = Hu$
where
\begin{eqnarray*}
z =   \left[ \begin{array}{c}
      x_{t} \\
      y_{t}
      \end{array} \right],
H =   \left[ \begin{array}{cc}
      I & 0 \\
      C & I
      \end{array} \right]
u =   \left[ \begin{array}{c}
      x_{t} \\
      w_t
      \end{array} \right],      
\end{eqnarray*}
Now we need to take the covariance of the simpler equation:
\begin{eqnarray*}
cov(z) &=& H cov(u) H'\\
cov(u) &=& \left[ \begin{array}{cc}
      P_{t} & 0 \\
      0 & R
      \end{array} \right]     
\end{eqnarray*}
The full product is:
\begin{eqnarray*}
\left[ \begin{array}{cc}
      I & 0 \\
      C & I
      \end{array} \right]
\left[ \begin{array}{cc}
      P_{t} & 0 \\
      0 & R
      \end{array} \right]
\left[ \begin{array}{cc}
      I & C' \\
      0 & I
      \end{array} \right]           
\end{eqnarray*}
which results in
\begin{eqnarray*}
\left[ \begin{array}{cc}
      P_{t} & P_{t}C' \\
      CP_{t} & CP_{t}C'+R
      \end{array} \right]           
\end{eqnarray*}
We can also re-write this for conditional and with the mean;
\begin{eqnarray}
\left[ \begin{array}{c}
      \hat{x}_{t}^t \\
      C\hat{x}_{t}^t
      \end{array} \right]
\:and\:      
\left[ \begin{array}{cc}
      P_{t}^t & P_{t}^tC' \\
      CP_{t}^t & CP_{t}^tC'+R
      \end{array} \right] \mlabel{3}
\end{eqnarray}
Same for the joint distribution for $x_{t+1}, y_{t+1}$.
\begin{eqnarray}
\left[ \begin{array}{c}
      \hat{x}_{t+1}^t \\
      C\hat{x}_{t+1}^t
      \end{array} \right]
\:and\:      
\left[ \begin{array}{cc}
      P_{t+1}^t & P_{t+1}^tC' \\
      CP_{t+1}^t & CP_{t+1}^tC'+R
      \end{array} \right] \mlabel{4}
\end{eqnarray}

Now in order to get statements for mean and variance for $x_{t+1}^{t+1}$, we
need to understand partitioned Gaussians. This will guide us when we use
portions of (4) and deriving the final equation \cite{jordan}.

@@

\begin{eqnarray}
\mu = 
\left[ \begin{array}{c}
      \mu_1 \\
      \mu_2
      \end{array} \right]           
\Sigma = 
\left[ \begin{array}{cc}
      \Sigma_{11} & \Sigma_{12}\\
      \Sigma_{21} & \Sigma_{22}
      \end{array} \right] \mlabel{6}
\end{eqnarray}
\begin{eqnarray*}
p(x|\mu,\Sigma) &=& \frac{1}{(2\pi)^{(p+q)/2}|\Sigma|^{1/2}} \\
&& exp\bigg\{
-\frac{1}{2}
\left( \begin{array}{c}
      x_1 - \mu_1 \\
      x_2 - \mu_2
      \end{array} \right)'
\bigg\}
\left[ \begin{array}{cc}
      \Sigma_{11} & \Sigma_{12}\\
      \Sigma_{21} & \Sigma_{22}
      \end{array} \right]^{-1}
\left( \begin{array}{c}
      x_1 - \mu_1 \\
      x_2 - \mu_2
      \end{array} \right)
\bigg\} 
\end{eqnarray*}
After much algebra, we can obtain equation for $p(x_1|x_2)$. From this, we can
get {\em conditioned} $\mu$ and $\Sigma$, as
\begin{eqnarray}
\mu_{1|2} &=& \mu_1 + \Sigma_{12}\Sigma_{22}^{-1}(x_2 - \mu_2) \mlabel{5} \\
\Sigma_{1|2} &=& \Sigma_{11}-\Sigma_{12}\Sigma_{22}^{-1}\Sigma_{21} \nonumber
\end{eqnarray}
Now plugging elements from  (5) into (4) according
to placement in (6) we can formulize $\hat{x}_{t+1}^{t+1}$ and
$P_{t+1}^{t+1}$. 
\begin{eqnarray*}
\hat{x}_{t+1}^{t+1} &=& \hat{x}_{t+1}^t + P_{t+1}^tC'
                  (CP_{t+1}^tC'+ \Sigma_{w})^{-1} \\
               &&   (y_{t+1} - C\hat{x}_{t+1}^t) \\
P_{t+1}^{t+1} &=& P_{t+1}^t - P_{t+1}^tC' (CP_{t+1}^tC' + R)^{-1}CP_{t+1}^t
\end{eqnarray*}
If we declare $K_t \equiv P_{t+1}^tC' (CP_{t+1}^tC'+ \Sigma_{w})^{-1}$ then,
\begin{eqnarray}
\hat{x}_{t+1}^{t+1} &=& \hat{x}_{t+1}^t + K_t (y_{t+1} -
C\hat{x}_{t+1}^t) \label{xtplus1} \\
P_{t+1}^{t+1} &=& P_{t+1}^t - K_t C P_{t+1}^t
\end{eqnarray}

@@

Based on (3), we can write the covariance of joint distribution
$(x_t,x_{t+1})$. In (3) the multiplication was with $C$, here it's
$A$. Hence,
\begin{eqnarray*}
\left[ \begin{array}{cc}
      P_{t}^t & P_{t}^tA' \\
      AP_{t}^t & AP_{t}^tA'+R
      \end{array} \right]
\end{eqnarray*}
Combining the previous equations together, we need to reach to a statement for
$x_{t}'$. We can start with $\mathbf{E}[x_t|x_{t+1},y_0,...,y_t)]$.
\begin{eqnarray*}
\mu_{1|2} &=& \mu_1 + \Sigma_{12}\Sigma_{22}^{-1}(x_2 - \mu_2)
\end{eqnarray*}
which becomes
\begin{eqnarray*}
\mathbf{E}[x_t|x_{t+1},y_0,...,y_t)] = \hat{x}_{t}^t+P_{t}^tA'(P_{t+1}^t)^{-1}(x_{t+1}-\hat{x}_{t+1}^t)
\end{eqnarray*}
Define $L_t \equiv P_{t}^tA'(P_{t+1}^t)^{-1}$. Then,
\begin{eqnarray*}
\mathbf{E}[x_t|x_{t+1},y_0,...,y_t)] =
\hat{x}_{t}^t+L_t(x_{t+1}-\hat{x}_{t+1}^t)
\end{eqnarray*}
How about $Var[x_t|x_{t+1},y_0,...,y_t]$? If we remember the joint covariance
for $(x_t,x_{t+1})$
\begin{eqnarray*}
 \left[ \begin{array}{cc} 
      P_{t}^t & P_{t}^tA' \\
      AP_{t}^t & AP_{t}^tA'+R
      \end{array} \right] && \\
\end{eqnarray*}
Then
\begin{eqnarray*}
\Sigma_{1|2} &=& \Sigma_{11}-\Sigma_{12}\Sigma_{22}^{-1}\Sigma_{21}
\end{eqnarray*}
becomes
\begin{eqnarray*}
Var[x_t|x_{t+1},y_0,...,y_t] &=& P_{t}^t - P_{t}^tA'(P_{t+1}^t)^{-1}AP_{t}^t
\end{eqnarray*}
Why?
\begin{eqnarray*}
\left[ \begin{array}{cc}
      P_{t}^t & P_{t}^tA' \\
      AP_{t}^t & AP_{t}^tA'+R
      \end{array} \right]
=      
\left[ \begin{array}{cc}
      P_{t}^t & P_{t}^tA' \\
      AP_{t}^t & P_{t+1}^t
      \end{array} \right]
\end{eqnarray*}
If we try to substitute $L_t \equiv P_{t}^tA'(P_{t+1}^t)^{-1}$ in to the
equation:
\begin{eqnarray*}
Var[x_t|x_{t+1},y_0,...,y_t] &=& P_{t}^t - P_{t}^t A'(P_{t+1}^t)^{-1}AP_{t}^t \\
&=& P_{t}^t - P_{t}^tA'(P_{t+1}^t)^{-1} (P_{t+1}^t(P_{t+1}^t)^{-1}) AP_{t}^t\\
&=& P_{t}^t - (P_{t}^tA'(P_{t+1}^t)^{-1}) P_{t+1}^t ((P_{t+1}^t)^{-1} AP_{t}^t)\\
&=& P_{t}^t - L_t  P_{t+1}^t L_t'
\end{eqnarray*}
Due to conditional independence, when we condition on $x_{t+1}$, we make $x_t$
independent of future observations $y_{t+1},...,y_T$. Hence we can simply say;
\begin{eqnarray*}
E[x_t|x_{t+1},y_0,...,y_T]  &=& E[x_t|x_{t+1},y_0,...,y_t]\\
&=& \hat{x}_{t}^t +L_t(x_{t+1}-\hat{x}_{t+1}^t) \\
Var[x_t|x_{t+1},y_0,...,y_T]  &=& Var[x_t|x_{t+1},y_0,...,y_t] \\
&=& P_{t}^t - L_t  P_{t+1}^t L_t'
\end{eqnarray*}
We are almost done. If we can drop the $x_{t+1}$ from LHS of both equations, we
can have our smoothing equations. Due to the so-called Tower Property:
\begin{eqnarray*}
\mathbf{E}[X|Z] &=& \mathbf{E}[\mathbf{E}[X|Y,Z]|Z]\\
Var[X|Z] &=& Var[\mathbf{E}[X|Y,Z]|Z] + \mathbf{E}[Var[X|Y,Z]|Z]
\end{eqnarray*}
we can drop $x_{t+1}$ from LHS.
\begin{eqnarray}
\hat{x}_{t}' & \equiv & \mathbf{E}[x_t|y_0,...,y_T] \nonumber\\
& = & \mathbf{E}[\mathbf{E}[x_t|x_{t+1},y_0,...,y_T]|y_0,...,y_T] \nonumber\\
& = & \mathbf{E}[\hat{x}_{t}^t +L_t(x_{t+1}-\hat{x}_{t+1}^t)] \nonumber\\
& = & \hat{x}_{t}^t +L_t(\hat{x}_{t+1}' -\hat{x}_{t+1}^t) \label{xsmooth}
\end{eqnarray}
Using similar algebra, we get
\begin{eqnarray*}
P_{t}'  = P_{t}^t + L_t(P_{t+1}'-P_{t+1}^t)L_t'
\end{eqnarray*}
